The role of cognateness in non-native spoken word recognition

### 2024-05-09 manuscript/appendix.qmd
Add Methods section to Experiment 3

---

---

# Appendix 1: MCMC diagnostics

```{r appendix-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

suppressWarnings({
  bayesplot_theme_set(
    new = theme_custom() + 
      theme(
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8),
        strip.text = element_text(size = 10),
        legend.key = element_rect(colour = NA, fill = "white")
      )
  )
})

color_scheme_set("brightblue")
pars <- c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv")
pars_labels <- c(
  "Intercept",
  "Frequency (+1 SD)",
  "PTHN \u00d7 Levenshtein", 
  "Levenshtein (+1 SD)",
  "PTHN (+1 SD)"
)
```

## Pairs plot

```{r appendix-mcmc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
posterior_cp <- as.array(model_fits$fit_4)
np_cp <- nuts_params(model_fits$fit_4)

plot_pairs <- mcmc_pairs(
  posterior_cp, pars = pars,
  off_diag_args = list(size = 0.75),
  diag_args = list(facet_args = list(labeller = label_value(labels = pars_labels)))
)

plot_pairs
```

@@

```{r appendix-trace, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
(mcmc_trace(model_fits$fit_4, pars) + legend_none()) /
  mcmc_rank_overlay(model_fits$fit_4, pars = pars) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "top"
  )
```

## Gelman-Rubin convergence criterion (R-hat)

```{r appendix-rhat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
rhats <- rhat(model_fits$fit_4)
ratios <- neff_ratio(model_fits$fit_4)

mcmc_rhat_hist(rhats, binwidth = 0.00025, color = "white") + 
  mcmc_neff_hist(ratios, binwidth = 0.1) +
  plot_layout(guides = "keep", ncol = 1) &
  theme(
    legend.position = "top"
  )
```

>>>>

## Appendix 1: MCMC diagnostics

```{r fig-appendix-diagnostics}
#| label: fig-appendix-diagnostics
#| fig-width: 9
#| fig-height: 9
var_labels <- c("b_Intercept" = "Intercept",
                "b_freq_zipf_2" = "Frequency",
                "b_lv" = "Cognateness",
                "b_neigh_n_h" = "CLPN",
                "b_neigh_n_h:lv" = "Cognateness × CLPN",
                "b_group1" = "cat-ENG vs. spa-ENG")

list(exp_1_m0, exp_2_m0, exp_3_m1) |> 
    set_names(paste0("Experiment ", 1:3)) |> 
    map_dfr(gather_draws, `b_.*`, regex = TRUE, .id = "exp") |> 
    dplyr::filter(.variable != "b_Intercept") |> 
    mutate(.value = .value/4,
           .variable = factor(.variable, 
                              levels = names(var_labels),
                              labels = var_labels)) |> 
    ggplot(aes(.iteration, .value, colour = factor(.chain))) +
    facet_grid(.variable~exp, scales = "free") +
    geom_line() +
    labs(x = "Iteration",
         y = "Value",
         colour = "Chain",
         fill = "Chain",
         linetype = "Chain") +
    scale_colour_brewer(palette = "Reds") +
    theme(legend.position = "top",
          legend.box = "horizontal",
          panel.border = element_rect(fill = NA,
                                      colour = "grey"),
          legend.justification = c(1, 0),
          axis.title.x = element_blank(),
          axis.title.y = element_blank())
```

@@

```{r appendix-trace, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
(bayesplot::mcmc_trace(exp_1_m0, names(var_labels)) + bayesplot::legend_none()) /
  bayesplot::mcmc_rank_overlay(exp_1_m0, pars = names(var_labels)) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "top"
  )
```
